% !TeX spellcheck = en_US
\chapter{Introduction}%
\label{ch:introduction}%

This chapter motivates the problem of dexterous robotic grasping, states the specific challenges addressed in this work, summarizes the contributions, and outlines the structure of this thesis.

\section{Motivation}%
\label{sec:motivation}%

Multi-fingered dexterous hands offer robots the potential to interact with the world in a manner approaching human versatility.
Unlike parallel-jaw grippers, which are limited to a narrow set of grasp geometries, anthropomorphic hands can conform to arbitrary object shapes, reorient objects in-hand, and wield tools~\cite{Tang2024}.
Despite this potential, high \ac{DoF} hands remain largely confined to research laboratories.
The core bottleneck is not hardware, as several capable platforms exist, including the Allegro Hand, the Shadow Dexterous Hand, and the Agile Hand used in this work, but rather the difficulty of synthesizing reliable control policies for contact-rich tasks.

Traditional analytic approaches to grasp planning rely on known object models and precise contact modeling, assumptions that rarely hold in unstructured environments~\cite{Wu2022}.
Learning-based methods, particularly \ac{DRL}, have recently demonstrated impressive dexterous manipulation capabilities in simulation~\cite{Handa2024, Lin2025}.
However, transferring these policies to real hardware introduces a fundamental challenge: during training, RL agents have access to privileged information (exact object poses, contact forces, full state vectors) that is unavailable at deployment time, where the robot must rely on noisy sensory observations such as camera images.

The dominant strategy to bridge this gap is \emph{policy distillation}, in which a privileged \emph{teacher} policy is trained in simulation with full state access and its behavior is then distilled into a \emph{student} policy that operates from realistic sensor inputs~\cite{Singh2025, Lum2024}.
The DAgger algorithm~\cite{Ross2011} is widely used for this distillation step, as it addresses the distribution shift between expert demonstrations and student-induced trajectories through iterative data aggregation.
Yet standard DAgger is agnostic to safety: it queries the teacher uniformly, including in states where the student already performs well, and does not prevent the student from visiting dangerous states during training.
For dexterous hands, where unsafe actions can cause hardware damage (e.g., jamming or overloading finger joints) or lead to unstable contact configurations, a safety-aware distillation procedure is desirable.

\section{Problem Statement}%
\label{sec:problem_statement}%

From the situation outlined above, a general problem emerges: dexterous multi-fingered hands are hardware-ready, yet no scalable recipe exists for obtaining grasping policies that are simultaneously \emph{deployable on real hardware} (i.e., driven by realistic sensor inputs rather than privileged state) and \emph{safe to train}, in the sense that the training procedure itself does not rely on repeatedly visiting states that would damage the hand or the object on a physical system.

The currently dominant pipeline, in which a privileged \ac{RL} teacher is trained in simulation and subsequently distilled into a sensor-based student, addresses the first requirement but not the second.
Standard distillation algorithms treat the student purely as an imitation target and provide no mechanism to avoid unsafe states encountered during the student's own rollouts.
As dexterous grasping moves closer to real-world deployment, this gap becomes the limiting factor: the cost of an unsafe episode is no longer a reset in simulation but potential hardware damage, and the question of \emph{how} a student policy is trained becomes as important as the final task performance it achieves.

This thesis therefore addresses the problem of distilling a privileged dexterous grasping teacher into a vision-based student in a way that is both effective in terms of grasp success and explicit about training-time safety, and asks under what conditions such a safety-aware distillation can match or improve upon the standard approach.

\section{Contributions}%
\label{sec:contributions}%

The main contributions of this semester thesis are:

\begin{enumerate}
    \item \textbf{Dexterous grasping environment in Isaac Lab:} A direct RL environment (\texttt{DirectRLEnv}) in Isaac Lab for training arm-hand grasping policies with the Agile Hand and \ac{FR3}, including custom reward shaping and asset integration.
    \item \textbf{Privileged teacher policy:} A trained RL teacher policy that achieves robust dexterous grasping of diverse objects in simulation using privileged state information.
    \item \textbf{SafeDAgger-based distillation:} A comparative study of DAgger versus SafeDAgger for distilling the privileged teacher into a vision-based student, evaluating both task performance (\ac{GSR}) and training safety (number of unsafe episodes).
\end{enumerate}

This work lays the foundation for the subsequent master's thesis, which will extend the pipeline to include finger-gaiting, in-hand reorientation, and fine tool use such as screwdriving.

\section{Thesis Outline}%
\label{sec:thesis_outline}%

The remainder of this thesis is structured as follows:

\begin{description}
    \item[Chapter~\ref{ch:related_work}] reviews related work on dexterous grasping, policy distillation methods (DAgger, SafeDAgger), and sim-to-real transfer for manipulation.
    \item[Chapter~\ref{ch:methods}] formalizes the problem as a \ac{POMDP}, derives the privileged teacher training objective, and presents the mathematical framework for DAgger, SafeDAgger, and DexSafeDagger distillation.
    \item[Chapter~\ref{ch:experiments}] describes the experimental setup, including the hardware platform, simulation environment, reward design, and network architectures, presents teacher training results and SafeDAgger distillation experiments, and compares DAgger and SafeDAgger.
    \item[Chapter~\ref{ch:conclusion}] discusses the findings, limitations, and directions for future work in the master's thesis.
\end{description}
%
